An Alternative Geometric Interpretation of the Riemann-Stieltjes Integral

### 2017-11-08 paper.tex
cav intro update

\title{A 2D Geometric Interpretation of the Riemann-Stieltjes Integral}
\markright{Abbreviated Article Title}
\author{Woodrow Wilson and Herbert Hoover}

@@

\begin{abstract}
It is desirable to have geometric interpretations for difficult mathematical ideas. In this paper we present an intiutive geometric interpretation of the Riemann-Stieltjes integral. It can be 
summarized as follows: the Riemann-Stieltjes integral can be inerpreted as an infinite sum of initesimally small non-rectangular 
integration strips. This interpretation is surprisingly similar to the geometric intrepretation of the ordinary Riemann integral. 
It is therefore likely that this geometric interpretaion would narrow the logical rift between the Riemann integral and the Riemann-Stieltjes integral (i.e. make the 
Riemann-Stieltjes integral less abstract).
\end{abstract}

\noindent
Providing a geometrical interpretation for a mathematical concept can greatly improve ones understanding of it. The Riemann-Stieltjes 
integral is no exception. Bullock proposed a three dimensional geometric framework with which one can interpret the Riemann-Stieltjes 
integral. Pol proposes to interpret the Riemann-Stieltjes integral as the real distance covered by a moving object, whose speed was 
correctly recorded but at incorrect times. In this paper we present an intiutive two dimensional geometric interpretation of the Riemann-Stieltjes integral.\\

\noindent
?? utilized the idea of using non-rectangular integration strips to perform integration
and in doing so developed the Cavalieri integral. The area of a non-rectangular integration strip is computed via 
Cavalieri's principle.  also showed that a Cavalieri integral can easily be converted to a Riemann-Stieltjes integral. The 
reverse operation was however not addressed therein. In this paper we address this shortcomming. An automatic consequence of being 
able to convert a Riemann-Stieltjes integral to a Cavalieri integral is that a geometric interpretation of the Riemann-Stieltjes
becomes evident. This geometric interpretation is: the Riemann-Stieltjes integral can be interpreted as an infinite sum of infinitesimally small
non-rectangular integration strips. Interestingly enough, this interpretation of the Riemann-Stieltjes integral is very similar 
to the geometric interpretation of the Riemann integral. The difference being: in the case of the Riemann integral we employ 
rectangular instead of non-rectangular integration strips. This geometric interpretation has been quite useful while 
theaching the ordinary Riemann integral to undergraduate students. It, therefore, seems likely that undergraduate students would find the non-rectangular geometric interpretation
of the Riemann-Stieltjes integral just as useful. It is also interesting to mention that it has been proposed to first 
teach other integral definitions (which are easier to grasp) before moving on to the Riemann integral. It therefore seems plausible that
students will find it useful to first understand the Cavaleri integral before the Riemann-Stieltjes integral is taught to them.

@@

\begin{thebibliography}{1}
\bibitem{parker13} Adam Parker, Who solved the Bernoulli equation and how did they do it? \textit{Coll. Math. J.} \textbf{44} (2013) 89--97.

\bibitem{hopkins} Brian Hopkins, ed., \textit{Resources for Teaching Discrete Mathematics}, Mathematical Association of America, Washington DC, 2009.
\end{thebibliography}

>>>>

\title{An Alternative Geometric Interpretation of the Riemann-Stieltjes Integral}
\markright{Abbreviated Article Title}
\author{Woodrow Wilson and Herbert Hoover}

@@

\begin{abstract}
It is desirable to have geometric interpretations for difficult mathematical ideas. In this paper we present an intuitive geometric interpretation of the Riemann-Stieltjes integral. It can be 
summarized as follows: the Riemann-Stieltjes integral can be inerpreted as an infinite sum of infinitesimally small non-rectangular 
integration strips. This interpretation is surprisingly similar to the geometric interpretation of the ordinary Riemann integral. 
It is therefore likely that this geometric interpretation would reduce the large jump 
students experience when they first encounter the Riemann-Stieltjes integral (i.e. make the 
Riemann-Stieltjes integral less abstract).
\end{abstract}

\noindent
Providing a geometrical interpretation for a mathematical concept can greatly improve ones understanding of it. The Riemann-Stieltjes 
integral is no exception. G.~Bullock proposed a projection-based geometric framework with which one can interpret the Riemann-Stieltjes 
integral \cite{bullock1988}. Furthermore, I.~Podlubny proposed to interpret the Riemann-Stieltjes integral as the real distance covered by a moving object, whose speed was 
correctly recorded but at incorrect times \cite{podlubny2002}. In this paper we present an alternative, more initiative, geometric interpretation of the Riemann-Stieltjes integral.\\

\noindent
Ackermann et al. utilized the idea of using non-rectangular integration strips to perform integration
and in doing so developed the Cavalieri integral \cite{ackermann2012}. The area of a non-rectangular integration strip is computed via 
Cavalieri's principle \cite{andersen1985,eves1991, malik1984, wildberger2002, young1985}. They also showed that a Cavalieri integral can easily be converted to a Riemann-Stieltjes integral. The 
reverse operation was however not described in \cite{ackermann2012}. In this paper we address this shortcoming. This 
reverse operation actually allows us to contrive a novel geometric interpretation of the Riemann-Stieltjes integral.
This geometric interpretation is: the Riemann-Stieltjes integral can be interpreted as an infinite sum of infinitesimally small
non-rectangular integration strips. Interestingly enough, this interpretation of the Riemann-Stieltjes integral is very similar 
to the geometric interpretation of the Riemann integral. The difference being: in the case of the Riemann integral we employ 
rectangular instead of non-rectangular integration strips. This geometric interpretation has been quite useful while 
teaching the ordinary Riemann integral to undergraduate students \cite{jones2015,sealey2014}. It, therefore, seems likely that undergraduate students would find the non-rectangular geometric interpretation
of the Riemann-Stieltjes integral just as useful. It is also interesting to mention, that there is a few papers in literature that proposes to first 
teach simpler integral definitions (which are easier to grasp) before introducing students to the Riemann integral \cite{czarnocha2001,prabhu2008,ruffa2002,zaskis2014}. It therefore seems plausible that
students will find it useful to first understand the Cavalieri integral before the Riemann-Stieltjes integral is taught to them.

@@

\begin{thebibliography}{10}
\bibitem{ackermann2012} 
E.~R.~Ackermann, T.~L.~Grobler, W.~Kleynhans, J.~C.~Olivier, B.~P.~Salmon, A.~J.~van Zyl, \emph{Quaest. Math.} \textbf{35} (2012) 265--296.

\bibitem{andersen1985}
K.~Andersen, Cavalieri's Method of Indivisibles, \emph{Arch. Hist. Exact Sci.} \textbf{31} (1985) 291--367.

\bibitem{bartle1976}
R.~G Bartle, \emph{The Elements of Real Analysis}, New York: Wiley, 1976

\bibitem{bullock1988}
G.~L.~Bullock, A Geometric Interpretation of the Riemann-Stieltjes Integral, \emph{Amer. Math. Monthly} \textbf{95} (1988) 448--455.

\bibitem{czarnocha2001} B. Czarnocha, E. Dubinsky S. Loch, V. Prabhu, D. Vidakovic, Conceptions of Area: In Students and in History. \emph{College Math. J.} \textbf{32} (2001) 99--109.

\bibitem{eves1991}
H.~Eves, Two Surprising Theorems on Cavalieri Congruence, \emph{College Math. J.} \textbf{22} (1991) 118--124.

\bibitem{jones2015}
S.~R.~Jones, Areas, Anti-Derivatives, and Adding Up Pieces: Definite Integrals in Pure Mathematics and Applied Science Contexts, \emph{J. Math. Behav.}
\textbf{38} (2015) 9--28.

\bibitem{malik1984}
H.~A~Malik, A Note on Cavalieri Integration, \emph{Math. Mag.} \textbf{57} (1984) 154--156.

\bibitem{podlubny2002}
I.~Podlubny, Geometric and Physical Interpretation of Fractional Integration and Fractional Differentiation, \emph{Fract. Calc. Appl. Anal.} 
\textbf{5} (2002) 367--386.

\bibitem{prabhu2008}
V.~Prabhu, B. Czarnocha, Los Indivisibles en el C\'{a}lculo Contempor\'{a}neo, \emph{Ed. Mat.} \textbf{20} (2008) 53--88.

\bibitem{ruffa2002}
A.~A.~Ruffa, The Generalized Method of Exhaustion, \emph{Int. J. Math. Math. Sci.} \textbf{31} (2002) 345--351.

\bibitem{sealey2014}
V. Sealey, A framework for Characterizing Student Understanding of Riemann Sums and Definite Integrals, \emph{J. Math. Behav.}  \textbf{33} (2014) 230--245.

\bibitem{wildberger2002}
W.~N.~Wildeberger, A New Proof of Cavalieri's Quadrature Formula, \emph{Amer. Math. Monthly} \textbf{109} (2002) 843--845.

\bibitem{william1917}
E.~G.~William, Cavalieri's Theorem in his own Words, \emph{Amer. Math. Monthly} \textbf{95} (1917) 447--451.

\bibitem{young1985}
L. Lay-Young, S. Kangshen, The Chinese Concept of Cavalieri's Principle and Its Application, \emph{Hist. Math.}, \textbf{12} (1985) 219--228.

\bibitem{zaskis2014}
D.~Zaskis, C.~Rasmussen, S.~P.~Shen, A Mean-Based Approach for Teaching the Concept of Integration, \emph{Probl. Resour. Issues Math. Undergrad. Stud.} \textbf{24} (2014) 116--137.
\end{thebibliography}
